Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-03-07 Manuscript/manuscript.tex
Further edits incorportating connection with the current literature. There are still a few point to be done.

\usepackage[x11names]{xcolor}  % typesetting in color
%% Generate PDF/A-2u
%\usepackage[a-2u]{pdfx}

@@

\usepackage{achemso}  
\usepackage{rotating}  
%\usepackage{times} 
%\usepackage{lmodern}    %% Prefer Latin Modern fonts
\usepackage{graphicx} 
\usepackage{setspace} 
\usepackage{amsmath} 
\usepackage{epstopdf} 
\usepackage{csquotes} 
\usepackage{mhchem} 
\usepackage{chemfig} 
\usepackage[obeyFinal]{easy-todo}
\usepackage[english]{babel}
\usepackage{xr}
\externaldocument{manuscriptSUPPL}

@@

%% The hyperref package for clickable links in PDF and also for storing
%% metadata to PDF (including the table of contents).
%% Most settings are pre-set by the pdfx package.
%\hypersetup{unicode}
%\hypersetup{breaklinks=true}
%\hypersetup{colorlinks,allcolors=DodgerBlue4}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%% The manuscript does not need to include \maketitle, which is 
%% executed automatically. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{document}

@@

@@

The C--H bond order parameters were first averaged over the time for individual lipids and
then the standard error of the mean of the average over lipids was used as the error estimate
of the mean \cite{botan15,ollila16,nmrlipids_proj4}.

@@

The form factors from scattering and C-H bond order parameters from NMR experiments are good measures to validate
the lipid bilayer structure in MD simulations against experiments because they
are related to the bilayer dimensions (area per molecule and thickness) and
conformational fluctuations of individual lipids, respectively, and 
can be directly calculated from MD simulations~\cite{ollila16}.
The X-ray scattering form factors and C-H bond order parameters in
the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{pan14,nmrlipids_proj4}.
Here, we extend the available experimental dataset of POPS lipids by measuring the C--H bond
order parameters also from the acyl chain region using R-PDLF experiments as described previously \cite{ferreira13}.
We first determined the order parameters of each peak corresponding different C--H bonds detected
in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
Because the exact assignment of these peaks was not possible, we assigned the
acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
\todo{Tiago has to check and update this and figures in SI.}

To evaluate the effect of ECC on the structural properties of POPS lipid bilayer properties in MD simulations,
we compared the results from a simulation with original Lipid17 \cite{??} combined with the Dang ion parameters \cite{??}
with the results from ECC-POPS simulation and experiments (Fig. \ref{simVSexpNOions_POPS}).
with sodium counterions are prevously publihed
(or only counter ions)
from simulations and experiments. 
The experimental X-ray scattering form factors 
of a bilayer are well reproduced by ECC-POPS, 
while for Lipid17/Dang they are in a slight disagreement
(see Fig.~\ref{simVSexpNOions_POPS}). 
The agreement of Lipid17 can be improved by using different parametes for ions \cite{aqvist90},
which, however, bring other kinds of artifacts with them 
(e.g., incorrect pairing of ions, condensation at larger salt concentrations \cite{kohagen16, chen07, nmrlipids_proj4}).

@@

The interaction of various cations including calcium 
with  neutral and also negatively charged lipid bilayers
was thoroughly compared between simulations and experiments 
from the point of the electrometer concept \cite{roux90, seelig87, altenbach84}
in the references~\citenum{catte16, nmrlipids_proj4}. 
It was concluded that the interaction of calcium with zwitterionic PC bilayers 
is generally overestimated in all studied models \cite{catte16},
while none of the models was capable of interpreting the NMR structural parameters
with the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4}. 
Accounting for the electronic polarization was identified as 
a non-negligible correction to the interaction between cations and POPC bilayers in Ref.~\cite{melcr18}.

Here, we demonstrate that the same approach 
significantly improves the description of calcium binding 
also to the negatively charged PC:PS (5:1) bilayers. 
In correspondence with the Ref.~\citenum{nmrlipids_proj4},
we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
in PC and PS lipids 
to quantify the binding affinity of the calcium cations
and the induced structural changes
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
and to mark the improvement after treating electronic polarizability, 
we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
While the response of the order parameters of PC to \ce{CaCl2} concentrations
is qualitatively correct in the Lipid14/17 model \cite{lipid17},
it is quantitatively comparable to the experimental trends in ECC-lipids. 
Moreover, the structure of PS head group is fundamentaly improved
after accounting for the electronic polarization 
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).

@@

The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
are very challenging to interpret from both experiments \cite{roux90} 
and also from MD simulations \cite{nmrlipids_proj4}. 
The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
are not found to follow the experimental profiles, 
nor any common trend,
with the currently available models \cite{nmrlipids_proj4}. 
In Fig.~\ref{fig:delta_ordPar_CaCl_PCPS},
we plot these structural parameters for one of the models in that work, Lipid17, 
and for ECC-lipids. 
Although the agreement between ECC-lipids and experiments is not optimal,
the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
are well captured by the model. 
For example the order parameter $\beta$ in POPS 
adopts a saturated value in experiments and ECC-lipids,
but it is growing with the \ce{CaCl2} concentration in Lipid17.
In addition, 
the forking of the order parameters $\alpha$ 
is almost constant in the experiment and ECC-lipids,
but it varies greatly in th Lipid17.

@@

Such a largely diminished flexibility 
may arise from the binding of calcium to the \ce{COO^-} group in PS. 
Whe have measured the mean orientation of the \ce{COO^-} group 
as the connector of the carbon atoms 
forming the bond between the group and the $\beta$-carbon of the phospholipid 
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
The mean angle between the \ce{COO^-} vector and the membrane normal
changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
 at 620~mM buffer concentration of \ce{CaCl2}.
The conformational flexibility of the PS head group is hence restrained 
by the binding of calcium cations to the \ce{COO^-} group, 
which reorients towards the membrane interior and to the phosphate region, 
where calcium cations bind predominantly.

@@

Measurements using infrared spectroscopy 
combined with simulations with the CHARMM36 model 
with a correction for overestimated binding of calcium \cite{valentine18}
reveal similar trends for the populations at various PS moieties
suggesting that POPS interacts mostly through its \ce{COO^-} group. 
While this feature agrees nicely with our findings, 
the balance of the interactions of PS moieties with calcium cations may be suboptimal
and the interaction strength underestimated in total in those simulations \cite{nmrlipids_proj4}.

In contrast, the combination of NMR experiments 
and MD simulations with the same model, CHARMM36, but without the correction for calcium binding
also reveal long-lived interactions of \ce{Ca^{2+}} with \ce{PO4} groups. \cite{hallock18}
Although the ratios of the life times of coordination follow the same trends as in our work,
the lengths of simulation trajectories are insufficient 
to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, nmrlipids_proj4}). 
Moreover, the content of POPS in the bilayers in that work is
much higher above the threshold after which 
\ce{Ca^{2+}} cations induce lipid clustering. \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}
At such a high content of negatively charged lipids,
PS head groups are found to form even four-membered clusters \cite{hallock18}. 
Our simulations are, however, at different conditions,
in which the clustering of POPS does not happen, 
and PS lipids prefer 1:1 ratio with \ce{Ca^{2+}}
as plotted in Fig.~\ref{fig:cacl_complexes}
showing relative probabilities of \ce{Ca^{2+}} complexes with a certain number of PC or PS lipids. 
Compared to a neutral POPC bilayer
the negatively charged bilayer has its stoichiometry 
shifted towards complexes with three phospholipids and one calcium. 
This is obviously due to the presence of POPS, 
which also contributes to the pre-formed one- and two-membered clusters of POPC. 
Clusters of four or more lipids were not observed in either membrane.

\todo{ THE REMAINING PIECES TO BE DISCUSSED  --  consider wheter the dynamics slowdown or the dihedral distributions could make good contributions to this paper. }

@@

@@

\todo{SAMULI: I have putted here two preliminary sentences
JOSEF: I can write conclusions, but I would prefer to wait with that after your comments and edits. }
Applying ECC to POPS lipids improves cation binding affinity to membranes
mixed with POPC and also the response of PS headgroup to the bound cations.
However, further improvement of the parameters is needed to fully reproduce the
details of calcium binding to PS headgroups.

>>>>

\usepackage[x11names]{xcolor}  % typesetting in color
%% Generate PDF/A-2u
\usepackage[a-2u]{pdfx}

@@

\usepackage{achemso}  
\usepackage{rotating}  
%\usepackage{times} 
%\usepackage{lmodern}    %% Prefer Latin Modern fonts
\usepackage{graphicx} 
\usepackage{setspace} 
\usepackage{amsmath} 
\usepackage{epstopdf} 
\usepackage{csquotes} 
\usepackage{mhchem} 
\usepackage{chemfig} 
\usepackage[obeyFinal]{easy-todo}
\usepackage[english]{babel}

@@

%% The hyperref package for clickable links in PDF and also for storing
%% metadata to PDF (including the table of contents).
%% Most settings are pre-set by the pdfx package.
\hypersetup{unicode}
\hypersetup{breaklinks=true}
\hypersetup{colorlinks,allcolors=DodgerBlue4}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%% The manuscript does not need to include \maketitle, which is 
%% executed automatically. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{document}

@@

To validate the structures of lipid bilayers, the scattering form factors and C-H bond order parameters
were calculated and compared against the experimental data as described previously~\cite{botan15,ollila16,melcr18,nmrlipids_proj4}.
%in simulations without ions were validated against NMR 
%by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating 
%as was done in Refs.~\citenum{catte16, melcr18, nmrlipids_proj4}.
%NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. 
%The simulated order parameters were calculated for all C-H bonds in lipid molecules, 
%while all order paremeters in the head group and only some in the \ce{C-H} groups in tails were resolved experimentally
%(see next section and Fig.~\ref{simVSexpNOions_POPS}). 
The order parameters for each C-H bond of individual lipids in the trajectory were first averaged
over the time. Then the average over lipids was taken and the standard error of the mean of this
average was used as the error estimate of the mean \cite{botan15,ollila16,nmrlipids_proj4}.

@@

@@

We compared X-ray scattering form factors and NMR order parameters of bilayers
in pure water without any ions (or only counter ions)
from simulations and experiments. 
The experimental X-ray scattering form factors 
of a bilayer are well reproduced by ECC-POPS, 
while for Lipid17/Dang they are in a slight disagreement
(see Fig.~\ref{simVSexpNOions_POPS}). 
The agreement of Lipid17 can be improved by using different parametes for ions \cite{aqvist90},
which, however, bring other kinds of artifacts with them 
(e.g., incorrect pairing of ions, condensation at larger salt concentrations \cite{kohagen16, chen07, nmrlipids_proj4}).

@@

The interaction of various cations including calcium 
with  neutral and also negatively charged lipid bilayers
was thoroughly compared between simulations and experiments 
from the point of the electrometer concept \cite{roux90, seelig87, altenbach84}
in the references~\citenum{catte16, nmrlipids_proj4}. 
It was concluded that the interaction of calcium with zwitterionic PC bilayers 
is generally overestimated in all studied models \cite{catte16},
while none of the models was capable of interpreting the NMR structural parameters
with the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4}. 
Accounting for the electronic polarization was identified as 
a non-negligible correction to the interaction between cations and POPC bilayers in Ref.~\cite{melcr18}. 
Here, we demonstrate that the same approach 
significantly improves the description of calcium binding 
also to the negatively charged PC:PS (5:1) bilayers. 
In correspondence with the Ref.~\citenum{nmrlipids_proj4},
we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
in PC and PS lipids 
to quantify the binding affinity of the calcium cations
and the induced structural changes
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
and to mark the improvement after treating electronic polarizability, 
we show simulation results from ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
While the response of the order parameters of PC to \ce{CaCl2} concentrations
is qualitatively correct in the Lipid14/17 model \cite{lipid17},
it is quantitatively comparable to the experimental trends in ECC-lipids. 
Moreover, the structure of PS head group is fundamentaly improved
after accounting for the electronic polarization 
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).

@@

The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
are very challenging to interpret from both experiments \cite{roux90} 
and also from MD simulations \cite{nmrlipids_proj4}. 
The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
are not found to follow the experimental profiles, 
nor any common trend,
with the currently available models \cite{nmrlipids_proj4}. 
In Fig.~\ref{fig:delta_ordPar_CaCl_PCPS},
we plot these structural parameters for one of the models in that work, Lipid17, 
and for ECC-lipids. 
Although the agreement between ECC-lipids and experiments is far from optimal,
the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
are well captured by the model. 
For example the order parameter $\beta$ in POPS 
adopts a saturated value in experiments and ECC-lipids,
but it is growing with the \ce{CaCl2} concentration in Lipid17.
In addition, 
the forking of the order parameters $\alpha$ 
is almost constant in the experiment and ECC-lipids,
but it varies greatly in th Lipid17.

@@

Such a largely diminished flexibility 
may arise from the binding of calcium to the \ce{COO^-} group in PS. 
Whe have measured the mean orientation of the \ce{COO^-} group 
as the connector of the carbon atoms 
forming the bond between the group and the $\beta$-carbon of the phospholipid. 
The mean angle between the \ce{COO^-} vector and the membrane normal
changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
 at 620~mM buffer concentration of \ce{CaCl2}.
The conformational flexibility of the PS head group is hence restrained 
by the binding of calcium cations to the \ce{COO^-} group, 
which reorients towards the membrane interior and to the phosphate region, 
where calcium cations bind predominantly.

@@

Recent simulations reveal similar trends for the populations at various PS moieties. \cite{valentine18, hallock18}
The results from infrared spectroscopy combined with 
simulations with the CHARMM36 model 
including a correction for overestimated binding of calcium \cite{valentine18}
suggest that POPS interacts mostly through it \ce{COO^-} group. 
In contrast, the combination of NMR experiments 
and MD simulations with the same model but without the correction 
also reveal long-lived interactions of \ce{Ca^{2+}} with \ce{PO4} groups. \cite{hallock18}
The content of POPS in the bilayers in that work is, however, 
much higher above the threshold after which 
\ce{Ca^{2+}} cations induce lipid clustering. \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}
At such highly charged membranes,
POPS lipids are found to form even four-membered clusters \cite{hallock18}. 
Our simulations are, however, at different conditions,
in which the clustering of POPS does not happen, 
and PS lipids prefer 1:1 ratio with \ce{Ca^{2+}}
as plotted in Fig.~\ref{fig:cacl_complexes}
showing relative probabilities of \ce{Ca^{2+}} complexes with a certain number of PC or PS lipids. 
Compared to a neutral POPC bilayer
the negatively charged bilayer has its stoichiometry 
shifted towards complexes with three phospholipids and one calcium. 
This is obviously due to the presence of POPS, 
which also contributes to the pre-formed one- and two-membered clusters of POPC. 
Clusters of four or more lipids were not observed in either membrane.

THE REMAINING PIECES TO BE DISCUSSED:

@@

%\todo{Stationary distribution: Make a figure documenting the populations of bound \ce{Ca^{2+}} cations (like I have in the presentation) that would accompany Table \ref{tab:Ca_binding_PCPS}. 
%This will roughly correspond to the PC stoichiometry plot \ref{fig:cacl_complexes}. }

@@

\todo{SAMULI: I have putted here two preliminary sentences}
Applying ECC to POPS lipids improves cation binding affinity to membranes
mixed with POPC and also the response of PS headgroup to the bound cations.
However, further improvement of the parameters is needed to fully reproduce the
details of calcium binding to PS headgroups.
